Parsing Postfix Log Files

### 2009-04-24 doc/logparser-background.tex
Changes made after proofreading the preface and chapters 1-3.

This work is part of a larger project to optimise a mail server's
Postfix-based anti-spam restrictions, generate statistics and graphs, and
provide a platform on which new restrictions can be evaluated to determine
if they are beneficial in the fight against spam.  The program written for
this project, \parsername{}, parses Postfix log files and populates a
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

Determining which Postfix restrictions reject the highest number of
delivery attempts is a short example of the optimisation possible using
data from the database:

\begin{verbatim}
SELECT name, description, restriction_name, hits_total
    FROM rules
    WHERE action = 'DELIVERY_REJECTED'
    ORDER BY hits_total DESC;
\end{verbatim}

@@

Another example is determining which restrictions are not effective:
\figureref{Sample SQL query showing ineffective restrictions} shows which
restrictions rejected fewer than 100 delivery attempts in the last log file
parsed, and the percentage of total rejections in that log file that each
of those restrictions represents.

@@

The sample database queries yield summary statistics about the efficiency of
anti-spam techniques.  These statistics are far less feasible to assess
directly from log files without prior pre-processing into a database in the
fashion proposed, implemented, and tested herein.

@@

\SMTPglossaryDescription{} Despite the simplicity of the protocol, many
virus and spam sending programs fail to implement it properly, so requiring
strict adherence to the protocol specification is beneficial in protecting
against spam and viruses.\footnote{Originally all mail servers adhered to
the principle of \textit{Be liberal in what you accept, and conservative in
what you send\/} from \acronym{RFC}~760~\cite{rfc760}, but unfortunately
that principle was written in a friendlier time.  Given the deluge of spam
that mail servers are subjected to daily, a more appropriate maxim could
be: \textit{Require strict adherence to relevant \acronyms{RFC}; implement
the strongest restrictions you can; relax the restrictions and adherence
only when legitimate mail is impeded.\/}  It is neither as friendly nor as
catchy, but it more accurately reflects the current circumstances.} A
typical \acronym{SMTP} conversation resembles the following (the lines
starting with a three digit number are sent by the server, all other lines
are sent by the client):

@@

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs have serious deficiencies; unfortunately, newer programs
tend to be written by competent programmers or send mail using well written
programs (e.g.\ Postfix or Sendmail on Unix hosts, Microsoft Outlook on
Windows hosts).  Traditionally a mail server would have done its best to
deal with deficient clients, with the intention of accepting as much mail
sent to its users as possible, e.g.\ by accepting sender or recipient
addresses that were not enclosed in \texttt{<>}, or by ignoring the absence
of a HELO command.\footnote{The HELO command is the first command sent by
the client in the \acronym{SMTP} conversation; it is required by the
protocol, but its original purpose was to detect a server trying to deliver
mail to itself, so some implementations omitted it.}  Given the volume of
spam sent daily, this liberal approach is no longer viable.

@@

Postfix is a \glsfirst{MTA} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive optional anti-spam restrictions,
allowing an administrator to employ those restrictions which they judge
suitable for their server's needs, rather than a fixed set chosen by
Postfix's author.  These restrictions can be selectively applied, combined,
and bypassed on a per-client, per-recipient, or per-sender basis, allowing
different levels of stricture and permissiveness.  Administrators can
supply their own rejection messages to make it clear to senders why their
mail was rejected.  Policy servers (\sectionref{policy servers}) provide a
simple way to write new restrictions without having to modify Postfix's
source code.  Unfortunately this flexibility has a cost: the complexity
Postfix's log files.  Although it is usually relatively simple to use
standard Unix text processing utilities to determine the fate of an
individual mail, sometimes it can be quite difficult.  For most mails the
journey is simple and brief, but the remainder can be quite complex (see
\sectionref{complications} for details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and it is
separated into multiple components that each perform one of the tasks
required of an \acronym{MTA}, e.g.\ receive mail, send mail, local delivery
of mail;\ full details can be found in~\cite{postfix-overview}.  Postfix's
design is strongly influenced by security concerns: those components that
interact with other hosts are not privileged,\footnote{Privilege means the
power to perform actions that are limited to the administrator, and not
available to ordinary users.} so bugs in those components will not give an
attacker extra privileges; those components that are privileged do not
interact with other hosts, making it much more difficult for an attacker to
exploit any bugs that may exist in those components.

@@

The restriction \texttt{check\_client\_access} checks if the IP address of
the connected client is found in the specified table and returns the
associated action if found; the method of searching the file is dependant
on the type of the lookup table~\cite{postfix-lookup-tables}.  Other
restrictions determine their result by consulting external sources,
e.g.\newline{} \tab{}\texttt{reject\_rbl\_client
dnsbl.example.com}\newline{} checks the \acronym{DNSBL}
\texttt{dnsbl.example.com} and rejects the command if the client's IP
address is listed.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno}, or the name of another restriction to be
evaluated; other less commonly used results are possible as described
in~\cite{access-man-page}.  The meaning of \textit{permit\/} and
\textit{reject\/} is obvious; \textit{dunno\/} means to stop evaluating the
current restriction and continue processing the remainder of the
restriction list, allowing exceptions to more general rules.  The
configuration example below shows how to require that all machines on the
local network except for the web server authenticate before sending mail;
the web server is exempt because the legacy applications running on it lack
authentication support.  The restriction list is evaluated from top to
bottom: \texttt{permit\_sasl\_authenticated} permits authenticated clients;
an exception is made for the web server in the
\texttt{check\_client\_access} check, and other local machines are
rejected, because they must be unauthenticated to reach that point in the
restriction list.

@@

That example also shows that it is trivial to supply a custom rejection
message.  When the result is the name of another restriction Postfix will
evaluate that restriction, allowing restrictions to be chosen based on the
client IP address, client hostname, HELO hostname, sender address, or
recipient address.  For example, the administrator may require that all
clients on the local network have valid DNS entries, to prevent people
sending mail from unknown machines; one example of how to achieve this is:

@@

The administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex user-defined sequences
of lookups, restrictions, and exceptions.  Postfix tries to protect the
administrator from misconfiguration in as far as is reasonable, e.g.\ the
restriction \texttt{check\_helo\_mx\_access} cannot cause a mail to be
accepted, because the parameter it checks (the hostname given in the HELO
command) is under the control of the remote client.  Despite this, it is
possible for the administrator to make catastrophic mistakes, e.g.\
rejecting all mail --- the administrator must be cognisant of the effects
their configuration changes will have.  This is similar to one of UNIX's
design philosophies: \textit{``UNIX was not designed to stop its users from
doing stupid things, as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

@@

A policy server~\cite{policy-servers} is an external program consulted by
Postfix to determine the fate of an \acronym{SMTP} command.  The policy
server is given state information by Postfix (sample state information is
shown in \tableref{Example attributes sent to policy servers}) and returns
a result (reject, permit, dunno, or a rejection name) as described in
\sectionref{Mixing and matching Postfix restrictions}.  A policy server can
perform more complex checks than those provided by Postfix: a trivial
example is allowing addresses associated with the payroll system to send
mail on the third Tuesday after pay day only, to help prevent problems from
phishing mails using faked sender addresses.  For example, a phishing mail
might pretend that the payroll system had a disastrous disk failure, and
until the server is replaced all salary payments will have to be processed
manually, so please reply to this mail with your name, address, and bank
account details; the criminal can then use any details sent to him to help
with identity theft.

@@

    \item \acronym{SPF} checks:
        \urlLastChecked{http://www.openspf.org/}{2009/04/21}.
        \acronym{SPF} records specify which mail servers are allowed to
        send mail using sender addresses from a particular domain.  The
        intention is to reduce spam from faked sender addresses,
        backscatter~\cite{postfix-backscatter}, and
        Joe~jobs~\glsadd{Joe-job}.  There has been considerable resistance
        to \acronym{SPF} because it breaks or vastly complicates some
        commonly-used features of \acronym{SMTP}, e.g.\ forwarding mail
        from one domain to another when a user moves.

    \item Greylisting,
        \urlLastChecked{http://www.greylisting.org/}{2009/04/21}, is a
        technique that temporarily rejects a delivery attempt when the
        triple of \newline{} \tab{}\texttt{(sender address, recipient
        address, remote IP address)}\newline{} has not been seen before; on
        second and subsequent delivery attempts from that triple the mail
        will be accepted.  This blocks spam from some senders because
        maintaining a list of failed addresses and retrying after a
        temporary failure is uneconomical for a spam sender, but a
        legitimate mail server must retry deliveries that fail temporarily.
        Sadly spam senders are using increasingly complex and well written
        programs to distribute spam, frequently using an ISP provided
        \acronym{SMTP} server from a compromised machine on the ISP's
        network.  Greylisting will slowly become less effective as spam
        senders adapt, but it does block a large percentage of spam mail at
        the moment; the most effective restrictions from the
        \numberOFlogFILES{} log files used when generating the results in
        \sectionref{Evaluation} are shown in \tableref{Most effective
        Postfix restrictions}.  The table shows that greylisting is worth
        using at the moment, particularly when you take into account its
        position as the final restriction that a mail must overcome in the
        configuration used on the mail server that generated the log files:
        on that server greylisting only takes effect for mails that have
        passed all other restrictions.  Some problems may be encountered
        when using greylisting: some servers fail to retry after a
        temporary failure, or legitimate mail may be delayed, particularly
        when coming from a pool of servers.

        \showtable{build/include-restriction-summary-table}{Most effective
        Postfix restrictions}{Most effective Postfix restrictions}

    \item Scoring systems such as postfwd,
        \urlLastChecked{http://postfwd.org/}{2009/04/21}, perform tests on
        features of the delivery attempt (e.g.\ IP address, sender
        address), incrementing or decrementing a score based on the
        results; if the final score is higher than a threshold the delivery
        attempt is rejected.  The administrator must manually whitelist
        clients if they are to bypass a Postfix restriction, whereas using
        a threshold that requires a delivery attempt to hit several scored
        restrictions will allow mail from clients that would be rejected by
        a binary restriction.

@@

This section has provided background information useful in understanding
this thesis, starting with the motivation behind the project, continuing
with an introduction to \acronym{SMTP}, and finishing with Postfix, its
anti-spam restrictions, and its support for policy servers.

>>>>

This work is part of a larger project to optimise a mail server's
Postfix-based anti-spam restrictions, generate statistics and graphs, and
provide a platform on which new restrictions can be developed and evaluated to determine
if they are beneficial in the fight against spam.  The program written for
this project, \parsername{}, parses Postfix log files and populates an
\acronym{SQL}
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

Determining the ten Postfix restrictions that reject the highest number of
delivery attempts is a short example of the analysis possible using data
from the database:

\begin{verbatim}
SELECT name, description, restriction_name, hits_total
    FROM rules
    WHERE action = 'DELIVERY_REJECTED'
    ORDER BY hits_total DESC
    LIMIT 10;
\end{verbatim}

@@

Another example is determining which restrictions are not effective: the
query in \figureref{Sample SQL query showing ineffective restrictions}
shows which restrictions rejected fewer than 100 delivery attempts in the
last log file parsed, and the percentage of total rejections in that log
file that each of those restrictions represents.

@@

The sample database queries yield summary statistics about the efficiency of
anti-spam techniques.  Analysis of this kind is much easier to perform when
the data source is a database that when trying to directly analyse log
files.

@@

\SMTPglossaryDescription{} Despite the simplicity of the protocol, many
viruses and spam sending programs fail to implement it properly, so
requiring strict adherence to the protocol specification is beneficial in
protecting against spam and viruses.\footnote{Originally all mail servers
adhered to the principle of \textit{Be liberal in what you accept, and
conservative in what you send\/} from \acronym{RFC}~760~\cite{rfc760}, but
unfortunately that principle was written in a friendlier time.  Given the
deluge of spam that mail servers are subjected to daily, a more appropriate
maxim could be: \textit{Require strict adherence to relevant
\acronyms{RFC}; implement the strongest restrictions you can; relax the
restrictions and adherence only when legitimate mail is impeded.\/}  It is
neither as friendly nor as catchy, but it more accurately reflects the
current circumstances.} A typical \acronym{SMTP} conversation resembles the
following (the lines starting with a three digit number are sent by the
server, all other lines are sent by the client):

@@

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs have serious deficiencies; unfortunately, newer programs
tend to be written by competent programmers or send mail using well written
programs (e.g.\ Postfix or Sendmail on Unix hosts, Microsoft Outlook on
Windows hosts).  Traditionally a mail server would have done its best to
deal with deficient clients, with the intention of accepting as much mail
sent to its users as possible, e.g.\ by accepting sender or recipient
addresses that were not enclosed in \texttt{<>}, or by ignoring the absence
of a HELO command.\footnote{The HELO command is the first command sent by
the client in the \acronym{SMTP} conversation; it is required by the
protocol, but its original purpose was to prevent mail loops by detecting a
server trying to deliver mail to itself, so some implementations omitted
it.}  Given the volume of spam sent daily, this liberal approach is no
longer viable.

@@

Postfix is a \acronym{MTA} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive optional anti-spam restrictions,
allowing an administrator to employ those restrictions which they judge
suitable for their server's needs, rather than a fixed set chosen by
Postfix's author.  These restrictions can be selectively applied, combined,
and bypassed on a per-client, per-recipient, or per-sender basis, allowing
different levels of stricture and permissiveness.  Administrators can
supply their own rejection messages to make it clear to senders why their
mail was rejected.  Policy servers (soon to be described in
\sectionref{policy servers}) provide a simple way to write new restrictions
without having to modify Postfix's source code.  Unfortunately this
flexibility has a cost: the complexity of Postfix's log files.  Although it
is usually relatively simple to use standard Unix text processing utilities
to determine the fate of an individual mail, sometimes it can be quite
difficult.  For most mails the journey through Postfix is simple and brief,
but the remainder can be quite complex (see \sectionref{complications} for
details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and it is
separated into multiple components that each perform one of the tasks
required of an \acronym{MTA}, e.g.\ receive mail, send mail, deliver mail
to a user's mailbox;\ full details can be found in~\cite{postfix-overview}.
Postfix's design is strongly influenced by security concerns: those
components that interact with other hosts are not
privileged,\footnote{Privilege means the power to perform actions that are
limited to the administrator, and are not available to ordinary users.} so
bugs in those components will not give an attacker extra privileges; those
components that are privileged do not interact with other hosts, making it
much more difficult for an attacker to exploit any bugs that may exist in
those components.

@@

The restriction \texttt{check\_client\_access} checks if the IP address of
the connected client is found in the specified table and returns the
associated result if found; the method of searching the file is dependant
on the type of the lookup table~\cite{postfix-lookup-tables}.  Other
restrictions determine their result by consulting external sources,
e.g.\newline{} \tab{}\texttt{reject\_rbl\_client
dnsbl.example.com}\newline{} checks the \acronym{DNSBL}
\texttt{dnsbl.example.com} and rejects the command if the client's IP
address is listed.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno}, or the name of another restriction to be
evaluated; other less commonly used results are possible as described
in~\cite{access-man-page}.  The meaning of \textit{permit\/} and
\textit{reject\/} is obvious; \textit{dunno\/} means to stop evaluating the
current restriction and continue processing the remainder of the
restriction list, allowing exceptions to more general rules.  The
configuration example below shows how to require that all machines on the
local network except for the web server authenticate before sending mail;
the web server is exempt because the legacy applications running on it lack
authentication support.  The restriction list is evaluated from top to
bottom: \texttt{permit\_sasl\_authenticated} permits authenticated clients;
an exception is made for the web server in the
\texttt{check\_client\_access} check, and then other local machines are
rejected, because to continue past \texttt{permit\_sasl\_authenticated} the
client must not have authenticated.

@@

That example also shows how to supply a custom rejection message.  When the
result of a lookup is the name of another restriction Postfix will evaluate
that restriction; this allows restrictions to be chosen based on the client
IP address, client hostname, HELO hostname, sender address, or recipient
address.  For example, the administrator may require that all clients on
the local network have valid DNS entries, to prevent people sending mail
from unknown machines; one example of how to achieve this is:

@@

The administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex user-defined sequences
of lookups, restrictions, and exceptions.  Postfix tries to protect the
administrator from misconfiguration in as far as it reasonably can, e.g.\
the restriction \texttt{check\_helo\_mx\_access} cannot cause a mail to be
accepted, because the parameter it checks (the hostname given in the HELO
command) is under the control of the remote client.  Despite this, it is
possible for the administrator to make catastrophic mistakes, e.g.\
rejecting all mail --- the administrator must be cognisant of the effects
their configuration changes will have.  This is similar to one of UNIX's
design philosophies: \textit{``UNIX was not designed to stop its users from
doing stupid things, as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

@@

A policy server~\cite{policy-servers} is an external program consulted by
Postfix to determine the fate of an \acronym{SMTP} command.  The policy
server is given state information by Postfix (sample state information is
shown in \tableref{Example attributes sent to policy servers}) and returns
a result (reject, permit, dunno, or a rejection name) as described in
\sectionref{Mixing and matching Postfix restrictions}.  A policy server can
perform more complex checks than those provided by Postfix, such as
allowing addresses associated with the payroll system to send mail on the
third Tuesday after pay day only, to help prevent problems from phishing
mails using faked sender addresses.  For example, a phishing mail might
pretend that the payroll system had a disastrous disk failure, and until
the server is replaced all salary payments will have to be processed
manually, so please reply to this mail with your name, address, and bank
account details; the criminal can then use any details sent to him to help
with identity theft.

@@

    \item Checking if the client satisfies a domain's \acronym{SPF}
        records, \urlLastChecked{http://www.openspf.org/}{2009/04/21}.
        \acronym{SPF} records specify which mail servers are allowed to
        send mail using sender addresses from a particular domain.  The
        intention is to reduce spam from faked sender addresses,
        backscatter, and joe~jobs~\glsadd{Joe-job}.  There has been
        considerable resistance to \acronym{SPF} because it breaks or
        vastly complicates some commonly-used features of \acronym{SMTP},
        e.g.\ forwarding mail from one domain to another when a user moves.

    \item Greylisting,
        \urlLastChecked{http://www.greylisting.org/}{2009/04/21}, is a
        technique that temporarily rejects a delivery attempt when the
        tuple of \newline{} \tab{}\texttt{(sender address, recipient
        address, remote IP address)}\newline{} has not been seen before; on
        second and subsequent delivery attempts from that tuple the mail
        will be accepted.  This blocks spam from some senders because
        maintaining a list of failed addresses and retrying after a
        temporary failure is uneconomical for a spam sender, but a
        legitimate mail server must retry deliveries that fail temporarily.
        Sadly spam senders are using increasingly complex and well written
        programs to distribute spam, frequently using an ISP provided
        \acronym{SMTP} server from a compromised machine on the ISP's
        network.  Greylisting will slowly become less effective as spam
        senders adapt, but it does block a large percentage of spam mail at
        the moment; the most effective restrictions from the
        \numberOFlogFILES{} log files used when generating the results in
        \sectionref{Evaluation} are shown in \tableref{Sample output from
        SQL query}.  That table shows that greylisting is worth using at
        the moment, particularly when you take into account its position as
        the final restriction that a mail must overcome in the
        configuration used on the mail server that generated the log files:
        on that server greylisting only takes effect for mails that have
        passed all other restrictions.  Some problems may be encountered
        when using greylisting: some servers fail to retry after a
        temporary failure, or legitimate mail may be delayed, particularly
        when coming from a pool of servers.

    \item Scoring systems such as postfwd,
        \urlLastChecked{http://postfwd.org/}{2009/04/21}, perform tests on
        features of the delivery attempt (e.g.\ IP address, sender
        address), incrementing or decrementing a score based on the
        results; if the final score is higher than a threshold the delivery
        attempt is rejected.  The administrator must manually whitelist
        clients if they are to bypass a Postfix restriction, whereas using
        a threshold that requires a delivery attempt to hit several scored
        restrictions will allow delivery attempt that would be rejected by
        a boolean restriction.

@@

This chapter has provided background information useful in understanding
this thesis, starting with the motivation behind the project, continuing
with an introduction to \acronym{SMTP}, and finishing with Postfix, its
anti-spam restrictions, and its support for policy servers.

### 2009-04-24 doc/logparser-implementation.tex
Changes made after proofreading the preface and chapters 1-3.

\chapter{Postfix Parser Implementation}

@@

\subsection{Actions In The \parsernamelong{}}

@@

\subsection{Example Rule}

@@

The example rule in \tableref{Example rule in implementation table}
recognises the log line that results from Postfix rejecting a delivery
attempt because the domain part of the sender address does not have an A,
AAAA, or MX DNS entry, i.e.\ mail could not be delivered to any address in
the sender's domain (for full details
see~\cite{reject-unknown-sender-domain}).  A sample log line that would be
recognised by this rule:

@@

% do not reformat this!
\begin{table}[thbp]
    \caption{Example of a rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

>>>>

\chapter{Postfix Log Parser}

@@

\subsection{Description Of Each Action}

@@

\subsection{Sample Rule}

@@

The sample rule in \tableref{Example rule in implementation table}
recognises the log line that results from Postfix rejecting a delivery
attempt because the domain part of the sender address does not have an A,
AAAA, or MX DNS entry, i.e.\ mail could not be delivered to any address in
the sender's domain (for full details
see~\cite{reject-unknown-sender-domain}).  A sample log line that would be
recognised by this rule:

@@

% do not reformat this!
\begin{table}[thbp]
    \caption{Sample rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

### 2009-04-24 doc/logparser-introduction.tex
Changes made after proofreading the preface and chapters 1-3.

The approach chosen to measure performance is to analyse the log files
produced by the \acronym{MTA} in use, Postfix, rather than modifying its
source code to generate statistics: this approach improves the chances of
other Postfix users testing and using the software developed for this
project.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix: the parser must be
designed and implemented so that adding support for parsing new inputs is a
simple task, because the log lines to be parsed will change over time.
This variety in log lines occurs for several reasons:

@@

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam techniques are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the first real foray the author, a Systems
Administrator for a network of over 2000 computers and over 1800 users,
took into processing Postfix log files.}  Very few will have performed
in-depth parsing and analysis of their log files, where the parsing must
correlate the log lines per-connection or per-queueid rather than
processing log lines independently.  One of the barriers to this kind of
processing is the unstructured nature of Postfix log files, where each log
line was added on an ad hoc basis when a requirement was discovered or new
functionality was added.\footnote{A history of all changes made to Postfix
is distributed with the source code, available from
\urlLastChecked{http://www.postfix.org/}{2009/02/23}.}  Further
complication arises because the set of log lines is not fixed, and log
lines can differ in many ways between servers, even within the same
organisation, where servers may be configured differently or running
different versions of Postfix.

@@

This architecture is the basis of \parsername{}, a program that parses
Postfix log files and places the resulting data into a database for later
analysis.  The gathered data can be used to optimise current anti-spam
defences, to produce statistics showing how effective each technique in use
is, or to provide a baseline to test new anti-spam measures against.
Numerous other uses are possible for such data: improving server
performance by identifying troublesome destinations and reconfiguring
appropriately; identifying regular high volume uses (e.g.\ customer
newsletters) and restricting those uses to off-peak times, or providing a
dedicated service for them; detecting virus outbreaks that propagate via
mail; as a base for billing customers on a shared server.  Preserving the
raw data enables a multitude of uses far beyond those conceived of by the
author.

@@

Chapter~\ref{Evaluation} evaluates \parsernames{} efficiency, discussing
the various optimisations implemented in the parser, and the effect they
have.  It also discusses the coverage achieved by \parsername{} over
\numberOFlogFILES{} log files, with separate sections for the number of log
lines correctly recognised, and the number of connections and mails whose
journey through Postfix was correctly reconstructed.

@@

Appendix~\ref{Postfix Daemons} provides a brief description of Postfix
daemons.

>>>>

The approach chosen to measure performance is to analyse the log files
produced by the \acronym{MTA} in use, Postfix, rather than modifying its
source code to generate statistics: this approach improves the chances of
other Postfix users testing, using, and benefiting from the software
developed for this project.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by Postfix: the
parser must be designed and implemented so that adding support for parsing
new inputs is a simple task, because the log lines to be parsed will change
over time.  This variety in log lines occurs for several reasons:

@@

Most mail server administrators will have performed some basic processing
of the log files produced by their mail server at one time or another,
whether it was to debug a problem, explain to a user why their mail is
being rejected, or check if new anti-spam techniques are working.  The more
adventurous will have generated statistics to show how successful each of
their anti-spam measures has been in the last week, and possibly even
generated some graphs to clearly illustrate these statistics to management
or users.\footnote{This was the first real foray the author, a Systems
Administrator for a network of over 2000 computers and over 1800 users,
took into processing Postfix log files.}  Very few will have performed
in-depth parsing and analysis of their log files, where the parsing must
correlate each connection or mail's log lines rather than
processing each log line in isolation.  One of the barriers to this kind of
processing is the unstructured nature of Postfix log files, where each log
line was added on an ad hoc basis when a requirement was discovered or new
functionality was added.\footnote{A history of all changes made to Postfix
is distributed with the source code, available from
\urlLastChecked{http://www.postfix.org/}{2009/02/23}.}  Further
complication arises because the set of log lines is not fixed, and log
lines can differ in many ways between servers, even within the same
organisation, where servers may be configured differently or running
different versions of Postfix.

@@

This architecture is the basis of \parsername{}, a program that parses
Postfix log files and places the resulting data into a \acronym{SQL}
database for later analysis.  The gathered data can be used to optimise
current anti-spam defences, to produce statistics showing how effective
each technique in use is, or to provide a baseline to test new anti-spam
measures against.  Numerous other uses are possible for such data:
improving server performance by identifying troublesome destinations and
reconfiguring appropriately; identifying regular high volume uses (e.g.\
customer newsletters) and restricting those uses to off-peak times or
providing a dedicated service for them; detecting virus outbreaks that
propagate via mail; billing customers on a shared server.  Preserving the
raw data enables a multitude of uses far beyond those conceived of by the
author.

@@

Chapter~\ref{Evaluation} evaluates \parsernames{} efficiency, exploring
the various optimisations implemented in the parser and the effect they
have.  It also discusses the coverage achieved by \parsername{} over
\numberOFlogFILES{} log files, with separate sections for the number of log
lines correctly recognised, and the number of connections and mails whose
journey through Postfix was correctly reconstructed.

@@

Appendix~\ref{Postfix Daemons} provides a brief description of each Postfix
daemon.
